---
title: "Phase 4 Data Exploration"
format: html
execute:
  echo: true
  warning: false
---

## Question and boundary

- **Question:** What cutoff-safe player-season rows and Phase 4 outputs are available in the canonical warehouse?
- **Data:** The same local DuckDB tables used by Python.
- **Unit:** One player and prediction season for features and targets; one candidate model prediction for evaluation.
- **Target:** Next-season fantasy points per active game, games active, or total fantasy points.
- **Cutoff:** Features for prediction season `s` stop before season `s` begins.
- **Validation:** This document audits stored rows; it does not select a model.
- **Caveat:** R is an optional read-only companion. Missing outputs are reported, not synthesized.

```{r}
library(DBI)
library(dplyr)
library(duckdb)
library(ggplot2)
library(knitr)
library(tibble)

find_project_root <- function(start = getwd()) {
  candidate <- normalizePath(start, winslash = "/", mustWork = TRUE)
  repeat {
    if (file.exists(file.path(candidate, "pyproject.toml"))) return(candidate)
    parent <- dirname(candidate)
    if (identical(parent, candidate)) stop("Could not find the repository root.")
    candidate <- parent
  }
}

project_root <- find_project_root()
warehouse_path <- file.path(project_root, "data", "warehouse", "fantasy_football.duckdb")
warehouse_ready <- file.exists(warehouse_path)
cat("Warehouse:", if (warehouse_ready) "found" else "not found", "\n")
```

```{r}
if (warehouse_ready) {
  con <- dbConnect(duckdb::duckdb(), dbdir = warehouse_path, read_only = TRUE)
  tables <- dbListTables(con)
  required <- c(
    "player_season_features", "player_season_targets", "baseline_predictions",
    "player_projection_runs", "player_projection_models",
    "player_projection_predictions", "player_projection_champions",
    "player_projection_board"
  )
  table_audit <- tibble(
    table = required,
    available = required %in% tables,
    rows = vapply(required, function(table_name) {
      if (!table_name %in% tables) return(NA_real_)
      as.numeric(dbGetQuery(con, paste0("SELECT count(*) AS rows FROM ", table_name))$rows[[1]])
    }, numeric(1))
  )
  kable(table_audit)
} else {
  cat("Initialize and build the canonical warehouse before running this audit.\n")
}
```

```{r}
if (warehouse_ready && all(c("player_season_features", "player_season_targets") %in% tables)) {
  coverage <- dbGetQuery(con, "
    SELECT
      f.prediction_season,
      f.position,
      count(*) AS feature_rows,
      count(t.player_id) AS target_rows,
      count(*) FILTER (
        WHERE CAST(json_extract(f.feature_payload, '$.is_rookie') AS BOOLEAN)
      ) AS rookie_feature_rows
    FROM player_season_features AS f
    LEFT JOIN player_season_targets AS t
      USING (player_id, prediction_season)
    GROUP BY f.prediction_season, f.position
    ORDER BY f.prediction_season, f.position
  ")
  kable(tail(coverage, 24))

  ggplot(coverage, aes(prediction_season, feature_rows, color = position)) +
    geom_line() +
    geom_point() +
    labs(
      title = "Cutoff-safe feature rows by prediction season",
      x = "Prediction season",
      y = "Feature rows",
      color = "Position"
    ) +
    theme_minimal()
}
```

```{r}
if (warehouse_ready) dbDisconnect(con, shutdown = TRUE)
```

The rookie column is a coverage diagnostic, not proof that historical rookie rows are safe for learned training. The Python feature audit and model status remain authoritative.
