---
title: "Regularized Regression Companion"
format: html
execute:
  echo: true
  warning: false
---

## Question and boundary

- **Question:** How does Ridge regularization stabilize a small position-specific player model?
- **Data:** Reviewed fields from canonical `player_season_features` joined to `player_season_targets`.
- **Unit:** One non-rookie WR and prediction season.
- **Target:** Next-season fantasy points per active game.
- **Cutoff:** Each feature payload stops before its prediction season.
- **Validation:** Tune on 2024 using only earlier prediction seasons; do not use 2025 to choose a penalty.
- **Caveat:** This reduced R example is educational and never replaces the production Python pipeline or champion decision.

```{r}
library(DBI)
library(dplyr)
library(duckdb)
library(ggplot2)
library(knitr)
library(parsnip)
library(purrr)
library(recipes)
library(tibble)
library(workflows)
library(yardstick)

find_project_root <- function(start = getwd()) {
  candidate <- normalizePath(start, winslash = "/", mustWork = TRUE)
  repeat {
    if (file.exists(file.path(candidate, "pyproject.toml"))) return(candidate)
    parent <- dirname(candidate)
    if (identical(parent, candidate)) stop("Could not find the repository root.")
    candidate <- parent
  }
}

warehouse_path <- file.path(
  find_project_root(), "data", "warehouse", "fantasy_football.duckdb"
)
warehouse_ready <- file.exists(warehouse_path)
```

```{r}
model_rows <- tibble()
if (warehouse_ready) {
  con <- dbConnect(duckdb::duckdb(), dbdir = warehouse_path, read_only = TRUE)
  tables <- dbListTables(con)
  if (all(c("player_season_features", "player_season_targets") %in% tables)) {
    model_rows <- dbGetQuery(con, "
      SELECT
        f.player_id,
        f.prediction_season,
        CAST(json_extract(f.feature_payload, '$.age_at_cutoff') AS DOUBLE) AS age_at_cutoff,
        CAST(json_extract(f.feature_payload, '$.weighted_3yr_fantasy_points_per_game') AS DOUBLE)
          AS weighted_3yr_fantasy_points_per_game,
        CAST(json_extract(f.feature_payload, '$.weighted_3yr_targets_per_game') AS DOUBLE)
          AS weighted_3yr_targets_per_game,
        CAST(json_extract(f.feature_payload, '$.history_seasons') AS DOUBLE) AS history_seasons,
        CAST(json_extract(t.target_payload, '$.fantasy_points_per_game') AS DOUBLE)
          AS fantasy_points_per_game
      FROM player_season_features AS f
      JOIN player_season_targets AS t USING (player_id, prediction_season)
      WHERE f.position = 'WR'
        AND NOT CAST(json_extract(f.feature_payload, '$.is_rookie') AS BOOLEAN)
        AND json_extract(t.target_payload, '$.fantasy_points_per_game') IS NOT NULL
        AND f.prediction_season <= 2024
      ORDER BY f.prediction_season, f.player_id
    ") |> as_tibble()
  }
}

if (nrow(model_rows) == 0) {
  cat("No eligible canonical rows are available; run the Python Phase 3 build first.\n")
} else {
  model_rows |> count(prediction_season) |> kable()
}
```

```{r}
ridge_results <- tibble()
if (nrow(model_rows) > 0 && any(model_rows$prediction_season == 2024)) {
  training <- model_rows |> filter(prediction_season < 2024)
  validation <- model_rows |> filter(prediction_season == 2024)

  ridge_recipe <- recipe(
    fantasy_points_per_game ~ age_at_cutoff +
      weighted_3yr_fantasy_points_per_game +
      weighted_3yr_targets_per_game + history_seasons,
    data = training
  ) |>
    step_impute_median(all_numeric_predictors()) |>
    step_normalize(all_numeric_predictors())

  penalties <- c(0.1, 1.0, 10.0)
  ridge_results <- map_dfr(penalties, function(value) {
    specification <- linear_reg(penalty = value, mixture = 0) |>
      set_engine("glmnet")
    fitted <- workflow() |>
      add_recipe(ridge_recipe) |>
      add_model(specification) |>
      fit(training)
    predictions <- predict(fitted, validation) |>
      bind_cols(validation |> select(fantasy_points_per_game))
    mae(predictions, truth = fantasy_points_per_game, estimate = .pred) |>
      mutate(penalty = value)
  })
  kable(ridge_results |> arrange(.estimate))
} else {
  cat("A 2024 validation slice is required for this educational fit.\n")
}
```

```{r}
if (exists("con")) dbDisconnect(con, shutdown = TRUE)
```

The table above is not the production result: it uses a deliberately small predictor subset to make the penalty visible. Production Python uses the complete locked feature contract, nested expanding folds, three targets, four positions, stored lineage, and matched baseline comparison.
