---
title: "Rolling Resampling Audit"
format: html
execute:
  echo: true
  warning: false
---

## Question and boundary

- **Question:** Do training seasons always precede the season being predicted?
- **Data:** Distinct canonical prediction seasons and stored Phase 4 candidate predictions.
- **Unit:** One expanding fold for the plan; one player/model/target prediction for the persisted audit.
- **Target:** Any Phase 4 player target.
- **Cutoff:** The latest training prediction season must be less than the evaluation season.
- **Validation:** Expanding 2020-2024 validation followed by the frozen 2025 test.
- **Caveat:** This document checks stored split metadata; Python owns the fit and all writes.

```{r}
library(DBI)
library(dplyr)
library(duckdb)
library(knitr)
library(rsample)
library(tibble)

find_project_root <- function(start = getwd()) {
  candidate <- normalizePath(start, winslash = "/", mustWork = TRUE)
  repeat {
    if (file.exists(file.path(candidate, "pyproject.toml"))) return(candidate)
    parent <- dirname(candidate)
    if (identical(parent, candidate)) stop("Could not find the repository root.")
    candidate <- parent
  }
}

warehouse_path <- file.path(
  find_project_root(), "data", "warehouse", "fantasy_football.duckdb"
)
warehouse_ready <- file.exists(warehouse_path)
```

```{r}
available_seasons <- integer()
stored_audit <- tibble()
if (warehouse_ready) {
  con <- dbConnect(duckdb::duckdb(), dbdir = warehouse_path, read_only = TRUE)
  tables <- dbListTables(con)
  if ("player_season_features" %in% tables) {
    available_seasons <- dbGetQuery(
      con,
      "SELECT DISTINCT prediction_season FROM player_season_features ORDER BY 1"
    )$prediction_season
  }
  if ("player_projection_predictions" %in% tables) {
    stored_audit <- dbGetQuery(con, "
      SELECT
        prediction_season,
        fold_label,
        min(training_max_season) AS min_training_max_season,
        max(training_max_season) AS max_training_max_season,
        count(*) AS prediction_rows,
        count(*) FILTER (WHERE training_max_season >= prediction_season) AS leakage_rows
      FROM player_projection_predictions
      WHERE actual_value IS NOT NULL
      GROUP BY prediction_season, fold_label
      ORDER BY prediction_season, fold_label
    ") |> as_tibble()
  }
}
```

```{r}
evaluation_seasons <- intersect(2020:2025, available_seasons)
fold_plan <- lapply(evaluation_seasons, function(season) {
  training <- available_seasons[available_seasons < season]
  tibble(
    evaluation_season = season,
    fold_label = if_else(season == 2025, "test", "validation"),
    training_seasons = paste(training, collapse = ", "),
    training_max_season = if (length(training)) max(training) else NA_integer_,
    passes_order_check = length(training) > 0 && max(training) < season
  )
}) |> bind_rows()

if (nrow(fold_plan) > 0) {
  kable(fold_plan)
} else {
  cat("No canonical seasons are available for a fold plan.\n")
}
```

```{r}
if (nrow(stored_audit) > 0) {
  stopifnot(all(stored_audit$leakage_rows == 0))
  kable(stored_audit)
} else {
  cat("No persisted Phase 4 predictions exist yet; run Python model training first.\n")
}
```

```{r}
if (exists("con")) dbDisconnect(con, shutdown = TRUE)
```

The validation labels above describe prediction seasons, not source-stat seasons. Any fitted imputation, scaling, encoding, tuning, or residual calibration must obey the same ordering.
