---
title: "Model Comparison and Calibration"
format: html
execute:
  echo: true
  warning: false
---

## Question and boundary

- **Question:** Which stored candidate won validation, how did the frozen test behave, and are its intervals empirically calibrated?
- **Data:** Canonical baseline predictions and Phase 4 learned predictions, champions, and interval fields.
- **Unit:** One player-season prediction by candidate and target.
- **Target:** Fantasy points per active game, games active, or total fantasy points.
- **Cutoff:** Every stored `training_max_season` must precede its prediction season.
- **Validation:** Pool 2020-2024 to select; display 2025 only after selection.
- **Caveat:** This is an independent read-only summary. The Python report and warehouse champion records are authoritative.

```{r}
library(DBI)
library(dplyr)
library(duckdb)
library(ggplot2)
library(knitr)
library(tibble)

find_project_root <- function(start = getwd()) {
  candidate <- normalizePath(start, winslash = "/", mustWork = TRUE)
  repeat {
    if (file.exists(file.path(candidate, "pyproject.toml"))) return(candidate)
    parent <- dirname(candidate)
    if (identical(parent, candidate)) stop("Could not find the repository root.")
    candidate <- parent
  }
}

warehouse_path <- file.path(
  find_project_root(), "data", "warehouse", "fantasy_football.duckdb"
)
warehouse_ready <- file.exists(warehouse_path)
learned <- tibble(
  player_id = character(), prediction_season = integer(), position = character(),
  target_name = character(), candidate_source = character(), candidate_name = character(),
  predicted_value = double(), actual_value = double(), p10 = double(), p50 = double(),
  p90 = double(), fold_label = character(), training_max_season = integer()
)
baselines <- tibble(
  player_id = character(), prediction_season = integer(), position = character(),
  target_name = character(), candidate_source = character(), candidate_name = character(),
  predicted_value = double(), actual_value = double(), fold_label = character()
)
champions <- tibble()
```

```{r}
if (warehouse_ready) {
  con <- dbConnect(duckdb::duckdb(), dbdir = warehouse_path, read_only = TRUE)
  tables <- dbListTables(con)
  if ("player_projection_predictions" %in% tables) {
    learned <- dbGetQuery(con, "
      SELECT
        player_id, prediction_season, position, target_name,
        'learned' AS candidate_source, model_family AS candidate_name,
        predicted_value, actual_value, p10, p50, p90, fold_label,
        training_max_season
      FROM player_projection_predictions
      WHERE actual_value IS NOT NULL
    ") |> as_tibble()
  }
  if ("baseline_predictions" %in% tables) {
    baselines <- dbGetQuery(con, "
      SELECT
        player_id, prediction_season, position, target_name,
        'baseline' AS candidate_source, baseline_name AS candidate_name,
        predicted_value, actual_value,
        CASE WHEN prediction_season = 2025 THEN 'test' ELSE 'validation' END AS fold_label
      FROM baseline_predictions
      WHERE actual_value IS NOT NULL AND prediction_season BETWEEN 2020 AND 2025
    ") |> as_tibble()
  }
  if ("player_projection_champions" %in% tables) {
    champions <- dbGetQuery(con, "
      SELECT * FROM player_projection_champions
      ORDER BY position, target_name
    ") |> as_tibble()
  }
}
```

```{r}
candidates <- bind_rows(
  learned |> select(
    player_id, prediction_season, position, target_name,
    candidate_source, candidate_name, predicted_value, actual_value, fold_label
  ),
  baselines
)

candidate_metrics <- candidates |>
  mutate(period = if_else(prediction_season == 2025, "test_2025", "validation_2020_2024")) |>
  filter(prediction_season %in% 2020:2025) |>
  group_by(position, target_name, candidate_source, candidate_name, period) |>
  summarise(
    rows = n(),
    mae = mean(abs(predicted_value - actual_value)),
    rmse = sqrt(mean((predicted_value - actual_value)^2)),
    .groups = "drop"
  )

if (nrow(candidate_metrics) > 0) {
  kable(candidate_metrics)
} else {
  cat("No matched Phase 4 candidate outputs exist yet; run Python model training first.\n")
}
```

```{r}
if (nrow(champions) > 0) {
  kable(champions)
} else {
  cat("No persisted champion records are available.\n")
}
```

```{r}
calibration <- learned |>
  filter(!is.na(p10), !is.na(p50), !is.na(p90)) |>
  group_by(position, target_name, candidate_name, fold_label) |>
  summarise(
    rows = n(),
    coverage_p10_p90 = mean(actual_value >= p10 & actual_value <= p90),
    mean_width_p10_p90 = mean(p90 - p10),
    pinball_p10 = mean(pmax(0.10 * (actual_value - p10), -0.90 * (actual_value - p10))),
    pinball_p50 = mean(pmax(0.50 * (actual_value - p50), -0.50 * (actual_value - p50))),
    pinball_p90 = mean(pmax(0.90 * (actual_value - p90), -0.10 * (actual_value - p90))),
    .groups = "drop"
  )

if (nrow(calibration) > 0) {
  kable(calibration)
  ggplot(calibration, aes(mean_width_p10_p90, coverage_p10_p90, color = position)) +
    geom_point() +
    geom_hline(yintercept = 0.80, linetype = "dashed") +
    labs(
      title = "Empirical interval coverage versus width",
      x = "Mean P10-P90 width",
      y = "Observed coverage",
      color = "Position"
    ) +
    theme_minimal()
} else {
  cat("No evaluable Phase 4 intervals are available.\n")
}
```

```{r}
if (exists("con")) dbDisconnect(con, shutdown = TRUE)
```

Do not promote a different candidate because its 2025 row looks better. On validation, lower learned MAE is still insufficient when the paired-bootstrap learned-minus-baseline interval crosses zero; that result is inconclusive and retains the baseline. Also inspect segment counts: an apparently strong metric on a small or mismatched sample is not a fair champion comparison.
